---
title: "ROCpAUC: Partial Area Indexes for ROC Curve Analysis"
subtitle: "A Bioconductor Package for High-Specificity and High-Sensitivity Biomarker Evaluation"
author: "Agustín Chumillas Roldán"
date: "`r Sys.Date()`"
output:
  BiocStyle::html_document:
    toc: yes
    toc_float: yes
    theme: spacelab
    number_sections: yes
    df_print: paged
    highlight: kate
    code_folding: show
  BiocStyle::pdf_document:
    toc: yes
    number_sections: yes
    latex_engine: pdflatex
    fig_caption: yes
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{ROCpAUC: Partial Area Indexes for ROC Curve Analysis}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo      = TRUE,
  message   = FALSE,
  warning   = FALSE,
  fig.align = "center",
  fig.width = 8,
  fig.height = 6,
  comment   = "#>"
)
```

# Introduction

The **ROCpAIv2** package provides a comprehensive framework for
evaluating the discriminatory performance of biomarkers and classifiers
via partial area under the ROC curve (pAUC) indexes.

The Receiver Operating Characteristic (ROC) curve plots the True
Positive Rate (TPR, sensitivity) against the False Positive Rate (FPR, 1
− specificity) at all classification thresholds. While the full AUC
summarises global performance, clinical practice often requires
restricting evaluation to a region of high specificity (low FPR) or high
sensitivity (high TPR). This package implements all four major scaled
partial-area indexes from the literature:

| Index | Domain | Condition for validity | Reference |
|---------------|---------------|-----------------------------|---------------|
| **MCpAUC** | High specificity (FPR) | Proper ROC curve (above the diagonal) | McClish (1989) |
| **TpAUC** | High specificity (FPR) | Any ROC curve shape | Vivo *et al.* (2018) |
| **NpAUC** | High sensitivity (TPR) | NLR-bounded curve | Jiang *et al.* (1996) |
| **FpAUC** | High sensitivity (TPR) | Any ROC curve shape | Franco & Vivo (2021) |

All main functions accept either a `data.frame` or a
`SummarizedExperiment` as input, return results as a `tibble` or
`SummarizedExperiment`, generate publication-quality ROC plots with
shaded partial-area regions, and provide bootstrap confidence intervals.

## Installation

```{r install, eval = FALSE}

# Install from Bioconductor
if (!requireNamespace("BiocManager", quietly = TRUE))
  install.packages("BiocManager")

BiocManager::install("ROCpAIv2")

library("ROCpAIv2")
```

------------------------------------------------------------------------

```{r include=FALSE}

library(dplyr)
library("ROCpAUC")

```

## Set Seed

```{r Set Seed}

# Set a random seed for reproducibility of all bootstrap and stochastic results
set.seed(1234567)

```

------------------------------------------------------------------------

# Load Example Dataset

```{r Load Sp Data}

# RNA-seq data for Schizosaccharomyces pombe (fission yeast) — main demo dataset
library(fission)

# Load the fission SummarizedExperiment (RNA-seq, S. pombe, wild-type vs mut7 strain)
data(fission)

# Select five genes of interest for all package demonstrations
genes_of_interest = c(
  "SPNCRNA.1080",   # non-coding RNA, used as a low-AUC control
  "SPAC186.08c",    # protein-coding gene
  "SPNCRNA.1420",   # non-coding RNA
  "SPCC70.08c",     # protein-coding gene
  "SPAC212.04c"     # protein-coding gene
)

# Build a plain data.frame version for demos that use data.frame input.
# Rows = samples, first column = binary strain label (wt vs mut7), rest = expression.
strain_vector = colData(fission)$strain
expr_matrix_t = t(assay(fission)[genes_of_interest, ])
Sp = as.data.frame(cbind(strain = strain_vector, expr_matrix_t))

# Inspect the first rows and verify column types
head(Sp)
col_classes = sapply(Sp, class)
table(col_classes)

```

------------------------------------------------------------------------

# Empirical ROC Curve — `pointsCurve`

Calculates the ROC curve coordinates — False Positive Rate (FPR) and
True Positive Rate (TPR) — from a binary reference vector and a
continuous or ordinal predictor. Optionally displays the ROC curve using
\code{ggplot2}.

## Demo

```{r demo-pointscurve-df}

# Input type 1: plain data.frame
# Plot the empirical ROC curve for the first gene using the pre-built Sp data.frame
pointsCurve(Sp[, 1], as.numeric(Sp[, 2]),
            plot  = TRUE,
            label = colnames(Sp)[2])

```

```{r demo-pointscurve-se}

# Input type 2: vectors extracted directly from a SummarizedExperiment
# Extract outcome and predictor from the SE object and pass them as vectors
outcome = SummarizedExperiment::colData(fission)$strain
expr    = SummarizedExperiment::assay(fission)["SPNCRNA.1080", ]

pointsCurve(as.numeric(outcome), as.numeric(expr),
            plot  = TRUE,
            label = "SPNCRNA.1080 (from SE)")

```

# Point-Estimate Functions {#point-estimate}

## pAUC — Raw Partial Area Under the ROC Curve

Returns the trapezoidal pAUC for a given FPR or TPR interval without any
scaling. Useful as a diagnostic building block or for direct comparison
with software that reports raw pAUC values. All four scaled indexes
(MCpAUC,TpAUC, NpAUC, FpAUC) are derived from this quantity

$$AUC = \int_0^1 TPR(FPR) \, d(FPR)$$

$$pAUC_{(t_1, t_2)} = \int_{t_1}^{t_2} TPR(FPR) \, d(FPR)$$

### Demo — pAUC

```{r demo-pauc-fpr-se}

# Input type 1: SummarizedExperiment, FPR domain [0, 0.25]
pAUC(fission, low.value = 0, up.value = 0.25,
     axis = "x", selection = genes_of_interest, variable = "strain")

```

```{r demo-pauc-tpr-se}

# Input type 2: SummarizedExperiment, TPR domain [0.9, 1]
pAUC(fission, low.value = 0.9, up.value = 1,
     axis = "y", selection = genes_of_interest, variable = "strain")

```

```{r demo-pauc-df}

# Input type 3: plain data.frame, FPR domain
pAUC(Sp, low.value = 0, up.value = 0.25, axis = "x")

```

------------------------------------------------------------------------

## MCpAUC — McClish Standardised pAUC

The **MCpAUC** index (McClish, 1989) rescales the raw partial area to
$[0.5, 1]$ using the diagonal as the minimum reference. Defined only for
*proper* ROC curves (above the diagonal throughout the region of
interest):

$$\widetilde{A}_{(t_1,t_2)} = \frac{1}{2}\!\left(1 + \frac{A_{(t_1,t_2)} - A^{\min}_{(t_1,t_2)}}{A^{\max}_{(t_1,t_2)} - A^{\min}_{(t_1,t_2)}}\right)$$

### Demo — MCpAUC

```{r demo-mcpauc-se-tibble}

# Input type 1: SummarizedExperiment → Output: tibble (with plots)
res_MC = MCpAUC(
  fission,
  low.value = 0, up.value = 0.25,
  selection = genes_of_interest, variable = "strain",
  plot = TRUE, plot_type = "both", plots_per_page = 5
)
res_MC

```

```{r demo-mcpauc-se-se}

# Input type 2: SummarizedExperiment → Output: SummarizedExperiment
se_MC = MCpAUC(
  fission,
  low.value = 0, up.value = 0.25,
  selection = genes_of_interest, variable = "strain",
  output_as_SE = TRUE
)
assay(se_MC)

```

```{r demo-mcpauc-df}

# Input type 3: plain data.frame → Output: tibble
MCpAUC(Sp, low.value = 0, up.value = 0.25)

```

------------------------------------------------------------------------

## TpAUC — Tighter Partial Area Index

The **TpAUC** (Vivo *et al.*, 2018) uses tighter bounds derived from the
actual TPR values at the interval endpoints, making it valid for both
proper and improper ROC curves and sensitive to crossing curves with
equal raw pAUC:

$$A^*_{(t_1,t_2)} = \frac{1}{2}\!\left(1 + \frac{A_{(t_1,t_2)} - A^*_{\min}}{A^*_{\max} - A^*_{\min}}\right)$$

### Demo — TpAUC

```{r demo-tpauc-se-tibble}

# Input type 1: SummarizedExperiment → Output: tibble (with plots)
res_Tp = TpAUC(
  fission,
  low.value = 0, up.value = 0.25,
  selection = genes_of_interest, variable = "strain",
  plot = TRUE, plot_type = "both", plots_per_page = 5
)
res_Tp

```

```{r demo-tpauc-se-se}

# Input type 2: SummarizedExperiment → Output: SummarizedExperiment
se_Tp = TpAUC(
  fission,
  low.value = 0, up.value = 0.25,
  selection = genes_of_interest, variable = "strain",
  output_as_SE = TRUE
)
assay(se_Tp)

```

```{r demo-tpauc-df}

# Input type 3: plain data.frame → Output: tibble
TpAUC(Sp, low.value = 0, up.value = 0.25)

```

------------------------------------------------------------------------

## NpAUC — Normalised pAUC (High-Sensitivity Domain)

**NpAUC** (Jiang *et al.*, 1996) summarises performance above a minimum
sensitivity $\text{TPR}_0$ by dividing the horizontal-band pAUC by the
width of that band:

$$N\widehat{pAUC}_{\text{TPR}_0} = \frac{A_{\text{TPR}_0}}{1 - \text{TPR}_0}$$

Valid only when the NLR is bounded; returns `NA` otherwise.

### Demo — NpAUC

```{r demo-npauc-se-tibble}

# Input type 1: SummarizedExperiment → Output: tibble (with plots)
res_Np = NpAUC(
  fission,
  low.value = 0.9, up.value = 1,
  selection = genes_of_interest, variable = "strain",
  plot = TRUE, plot_type = "both", plots_per_page = 5
)
res_Np

```

```{r demo-npauc-se-se}

# Input type 2: SummarizedExperiment → Output: SummarizedExperiment
se_Np = NpAUC(
  fission,
  low.value = 0.9, up.value = 1,
  selection = genes_of_interest, variable = "strain",
  output_as_SE = TRUE
)
assay(se_Np)

```

```{r demo-npauc-df}

# Input type 3: plain data.frame → Output: tibble
NpAUC(Sp, low.value = 0.9, up.value = 1)

```

------------------------------------------------------------------------

## FpAUC — Fitted Partial Area Index

**FpAUC** (Franco & Vivo, 2021) extends NpAUC using tighter NLR-based
bounds valid for *any* ROC curve shape. Always lies in $[0.5, 1]$ and
distinguishes crossing curves with equal raw pAUC.

| Case | NLR condition | FpAUC formula |
|-----------------|----------------------------|----------------------------|
| **Bounded** | $\text{NLR}(x) \leq \text{NLR}_0$ | $A_{TPR_0} \,/\, [(1-\text{FPR}_0)(1-\text{TPR}_0)]$ |
| **Partially proper** | $\text{NLR}(x) \leq 1$ | $[A_{TPR_0} + (1-\text{TPR}_0)(1-\text{FPR}_0)] \,/\, [2(1-\text{FPR}_0)(1-\text{TPR}_0)]$ |
| **Improper** | Some $\text{NLR}(x) > 1$ | $[A_{TPR_0} + (1+\text{TPR}_0)(\text{TPR}_0-\text{FPR}_0)] \,/\, [(1+\text{TPR}_0-2\text{FPR}_0)(1-\text{TPR}_0)]$ |

### Demo — FpAUC

```{r demo-fpauc-se-tibble}

# Input type 1: SummarizedExperiment → Output: tibble (with plots)
res_Fp = FpAUC(
  fission,
  low.value = 0.9, up.value = 1,
  selection = genes_of_interest, variable = "strain",
  plot = TRUE, plot_type = "both", plots_per_page = 5
)
res_Fp

```

```{r demo-fpauc-se-se}

# Input type 2: SummarizedExperiment → Output: SummarizedExperiment
se_Fp = FpAUC(
  fission,
  low.value = 0.9, up.value = 1,
  selection = genes_of_interest, variable = "strain",
  output_as_SE = TRUE
)
assay(se_Fp)

```

```{r demo-fpauc-df}

# Input type 3: plain data.frame → Output: tibble
FpAUC(Sp, low.value = 0.9, up.value = 1)

```

------------------------------------------------------------------------

## allIndices — All Four Indexes in One Call

Convenience wrapper that runs \code{MCpAUC}, \code{TpAUC}, \code{NpAUC},
and \code{FpAUC} and joins the results into a single wide \code{tibble}.
FPR-domain indexes (MCpAUC, TpAUC) use \code{fpr.low} / \code{fpr.up};
TPR-domain indexes (NpAUC, FpAUC) use \code{tpr.low} / \code{tpr.up}.
This is the recommended entry point for exploratory multi-index
comparison.

### Demo — allIndices

```{r demo-allindices-se}

# Input type 1: SummarizedExperiment → Output: tibble (with plots)
allIndices(
  fission,
  fpr.low = 0, fpr.up = 0.25,
  tpr.low = 0.9, tpr.up = 1,
  selection = genes_of_interest, variable = "strain",
  plot = TRUE, plot_type = "both", plots_per_page = 5
)

```

```{r demo-allindices-df}

# Input type 2: plain data.frame → Output: tibble
allIndices(
  Sp,
  fpr.low = 0, fpr.up = 0.25,
  tpr.low = 0.9, tpr.up = 1
)

```

------------------------------------------------------------------------

# Bootstrap Confidence Interval Functions {#bootstrap}

## MCpAUCboot

Estimates confidence intervals for the McClish standardised pAUC
(MCpAUC) via nonparametric bootstrap resampling. Variables for which
MCpAUC is undefined on the original sample (improper ROC curves) are
automatically skipped and return NA in all output columns.

### Demo — MCpAUCboot

```{r demo-mcpaucboot-se-tibble}

# Input type 1: SummarizedExperiment → Output: tibble (with diagnostic plots)
boot_MC = MCpAUCboot(
  fission,
  low.value = 0, up.value = 0.25,
  selection = genes_of_interest, variable = "strain",
  r = 100, plot = TRUE, plot_type = "both"
)
boot_MC

```

```{r demo-mcpaucboot-se-se}

# Input type 2: SummarizedExperiment → Output: SummarizedExperiment
boot_MC_se = MCpAUCboot(
  fission,
  low.value = 0, up.value = 0.25,
  selection = genes_of_interest, variable = "strain",
  r = 100, output_as_SE = TRUE
)
assay(boot_MC_se)

```

```{r demo-mcpaucboot-df}

# Input type 3: plain data.frame → Output: tibble
MCpAUCboot(Sp, low.value = 0, up.value = 0.25, r = 100)

```

------------------------------------------------------------------------

## TpAUCboot

Estimates confidence intervals for the tighter partial area index
(TpAUC) via nonparametric bootstrap resampling. Because TpAUC is valid
for any ROC curve shape, no variables are skipped (no NA from shape
conditions).

### Demo — TpAUCboot

```{r demo-tpaucboot-se}

# Input type 1: SummarizedExperiment → Output: tibble (with diagnostic plots)
boot_Tp = TpAUCboot(
  fission,
  low.value = 0, up.value = 0.25,
  selection = genes_of_interest, variable = "strain",
  r = 100, plot = TRUE, plot_type = "both"
)
boot_Tp

```

```{r demo-tpaucboot-df}

# Input type 2: plain data.frame → Output: tibble
TpAUCboot(Sp, low.value = 0, up.value = 0.25, r = 100)

```

------------------------------------------------------------------------

## NpAUCboot

Estimates confidence intervals for the normalised partial AUC (NpAUC)
via nonparametric bootstrap resampling. Variables for which NpAUC is
undefined (NLR condition not satisfied) on the original sample are
automatically skipped and return NA in all output columns. Consider
FpAUCboot for those variables.

### Demo — NpAUCboot

```{r demo-npaucboot-se}

# Input type 1: SummarizedExperiment → Output: tibble (with diagnostic plots)
boot_Np = NpAUCboot(
  fission,
  low.value = 0.9, up.value = 1,
  selection = genes_of_interest, variable = "strain",
  r = 100, plot = TRUE, plot_type = "both"
)
boot_Np

```

```{r demo-npaucboot-df}

# Input type 2: plain data.frame → Output: tibble
NpAUCboot(Sp, low.value = 0.9, up.value = 1, r = 100)

```

------------------------------------------------------------------------

## FpAUCboot

Estimates confidence intervals for the fitted partial area index (FpAUC)
via nonparametric bootstrap resampling. Because FpAUC is valid for any
ROC curve shape, no variables are skipped. This makes FpAUCboot the most
robust bootstrap function in the package, especially useful when
NpAUCboot returns NA for some variables.

### Demo — FpAUCboot

```{r demo-fpaucboot-se-tibble}

# Input type 1: SummarizedExperiment → Output: tibble (with diagnostic plots)
boot_Fp = FpAUCboot(
  fission,
  low.value = 0.9, up.value = 1,
  selection = genes_of_interest, variable = "strain",
  r = 100, plot = TRUE, plot_type = "both"
)
boot_Fp

```

```{r demo-fpaucboot-se-se}

# Input type 2: SummarizedExperiment → Output: SummarizedExperiment
boot_Fp_se = FpAUCboot(
  fission,
  low.value = 0.9, up.value = 1,
  selection = genes_of_interest, variable = "strain",
  r = 100, output_as_SE = TRUE
)
assay(boot_Fp_se)

```

```{r demo-fpaucboot-df}

# Input type 3: plain data.frame → Output: tibble
FpAUCboot(Sp, low.value = 0.9, up.value = 1, r = 100)

```

------------------------------------------------------------------------

# Comparative Summary Tables {#summary}

## Point-Estimate Indexes

Combine all four point-estimate tibbles side by side into one summary
table. pAUC_FPR = raw FPR-domain pAUC (basis for MCpAUC and TpAUC)
pAUC_TPR = raw TPR-domain pAUC (basis for NpAUC and FpAUC)

```{r summary-point-estimates}

summary_tbl = res_MC %>%
  dplyr::rename(pAUC_FPR = pAUC) %>%
  dplyr::left_join(
    dplyr::select(res_Tp, Variable, TpAUC),
    by = "Variable"
  ) %>%
  dplyr::left_join(
    dplyr::select(res_Np, Variable, NpAUC, pAUC_TPR = pAUC),
    by = "Variable"
  ) %>%
  dplyr::left_join(
    dplyr::select(res_Fp, Variable, FpAUC),
    by = "Variable"
  ) %>%
  dplyr::select(Variable, pAUC_FPR, MCpAUC, TpAUC, pAUC_TPR, NpAUC, FpAUC)

summary_tbl

```

------------------------------------------------------------------------

## Bootstrap Confidence Intervals

Build transposed bootstrap summary tables: genes as columns, metrics as
rows. This layout makes it easy to compare the same metric across genes
at a glance.

```{r summary-bootstrap}

# Build transposed bootstrap summary tables: genes as columns, metrics as rows.
# This layout makes it easy to compare the same metric across genes at a glance.

# Helper: pivot one bootstrap result to a metric × gene matrix
make_boot_matrix = function(boot_tbl, index_col) {
  # Reshape: one row per metric (estimate, bias, sd, lwr, upr), one column per gene
  metrics = c(index_col, "bias", "sd", "lwr", "upr")
  mat = t(as.matrix(boot_tbl[, metrics]))
  colnames(mat) = boot_tbl$Variable
  rownames(mat) = metrics
  return(as.data.frame(round(mat, 6)))
}

# FPR-domain bootstrap summary (MCpAUC and TpAUC) — genes as columns
cat("Bootstrap Summary — MCpAUC (FPR in [0, 0.25])\n")
make_boot_matrix(boot_MC, "MCpAUC")

```

```{r summary-bootstrap-Tp}

cat("Bootstrap Summary — TpAUC (FPR in [0, 0.25])\n")
make_boot_matrix(boot_Tp, "TpAUC")

```

```{r summary-bootstrap-Np}

cat("Bootstrap Summary — NpAUC (TPR in [0.9, 1])\n")
make_boot_matrix(boot_Np, "NpAUC")

```

```{r summary-bootstrap-Fp}

cat("Bootstrap Summary — FpAUC (TPR in [0.9, 1])\n")
make_boot_matrix(boot_Fp, "FpAUC")

```

------------------------------------------------------------------------

# SummarizedExperiment Output Demo {#se-demo}

Return MCpAUC results as a SummarizedExperiment for downstream
Bioconductor workflows (e.g., integration with DESeq2 or limma

```{r demo-se-assay}

se_MC = MCpAUC(
  fission,
  low.value = 0, up.value = 0.25,
  selection = genes_of_interest, variable = "strain",
  output_as_SE = TRUE
)

# The assay matrix: rows = metrics (MCpAUC and pAUC), columns = predictor variables
assay(se_MC)

```

```{r demo-se-rowdata}

# Row metadata: names of the computed metrics stored in the assay
rowData(se_MC)

```

```{r demo-se-coldata}

# Column metadata: names of the predictor variables
colData(se_MC)

```

------------------------------------------------------------------------

# References

Alon, U., Barkai, N., Notterman, D. A., Gish, K., Ybarra, S., Mack, D.,
& Levine, A. J. (1999). Broad patterns of gene expression revealed by
clustering analysis of tumor and normal colon tissues probed by
oligonucleotide arrays. *Proceedings of the National Academy of
Sciences*, *96*(12), 6745–6750.
<https://doi.org/10.1073/pnas.96.12.6745>

Bandos, A. I., Guo, B., & Gur, D. (2017). Estimating the area under ROC
curve when the fitted binormal curves demonstrate improper shape.
*Academic Radiology*, *24*(2), 209–219.
<https://doi.org/10.1016/j.acra.2016.09.020>

Canty, A., & Ripley, B. D. (2019). *boot: Bootstrap R (S-Plus)
functions* (R package version 1.3-24) [Software]. CRAN.
<https://CRAN.R-project.org/package=boot>

Davison, A. C., & Hinkley, D. V. (1997). *Bootstrap methods and their
application*. Cambridge University Press.
<https://doi.org/10.1017/CBO9780511802843>

Dudoit, S., Fridlyand, J., & Speed, T. P. (2002). Comparison of
discrimination methods for the classification of tumors using gene
expression data. *Journal of the American Statistical Association*,
*97*(457), 77–87. <https://doi.org/10.1198/016214502753479248>

Fawcett, T. (2006). An introduction to ROC analysis. *Pattern
Recognition Letters*, *27*(8), 861–874.
<https://doi.org/10.1016/j.patrec.2005.10.010>

Franco, M., & Vivo, J.-M. (2021). Evaluating the performances of
biomarkers over a restricted domain of high sensitivity. *Mathematics*,
*9*(21), 2826. <https://doi.org/10.3390/math9212826>

García, J.-P., Franco, M., & Vivo, J.-M. (2021). *ROCpAI: Receiver
Operating Characteristic Partial Area Indexes for evaluating
classifiers* [Bioconductor package].
<http://bioconductor.org/packages/ROCpAI/>

Golub, T. R., Slonim, D. K., Tamayo, P., Huard, C., Gaasenbeek, M.,
Mesirov, J. P., Coller, H., Loh, M. L., Downing, J. R., Caligiuri, M.
A., Bloomfield, C. D., & Lander, E. S. (1999). Molecular classification
of cancer: Class discovery and class prediction by gene expression
monitoring. *Science*, *286*(5439), 531–537.
<https://doi.org/10.1126/science.286.5439.531>

Greiner, M., Pfeiffer, D., & Smith, R. D. (2000). Principles and
practical application of the receiver-operating characteristic analysis
for diagnostic tests. *Preventive Veterinary Medicine*, *45*(1–2),
23–41.
[https://doi.org/10.1016/S0167-5877(00)00115-X](https://doi.org/10.1016/S0167-5877(00)00115-X){.uri}

Jiang, Y., Metz, C. E., & Nishikawa, R. M. (1996). A receiver operating
characteristic partial area index for highly sensitive diagnostic tests.
*Radiology*, *201*(3), 745–750.
<https://doi.org/10.1148/radiology.201.3.8939225>

Leong, H. S., Dawson, K., Wirth, C., Li, Y., Connolly, Y., Smith, D. L.,
Wilkinson, C. R. M., & Miller, C. J. (2014). A global non-coding RNA
system modulates fission yeast protein levels in response to stress.
*Nature Communications*, *5*, 3947. <https://doi.org/10.1038/ncomms4947>

Marguerat, S., Schmidt, A., Codlin, S., Chen, W., Aebersold, R., &
Bähler, J. (2012). Quantitative analysis of fission yeast transcriptomes
and proteomes in proliferating and quiescent cells. *Cell*, *151*(3),
671–683. <https://doi.org/10.1016/j.cell.2012.09.019>

McClish, D. K. (1989). Analyzing a portion of the ROC curve. *Medical
Decision Making*, *9*(3), 190–195.
<https://doi.org/10.1177/0272989X8900900307>

Metz, C. E. (1978). Basic principles of ROC analysis. *Seminars in
Nuclear Medicine*, *8*(4), 283–298.
[https://doi.org/10.1016/S0001-2998(78)80014-2](https://doi.org/10.1016/S0001-2998(78)80014-2){.uri}

Morgan, M., Obenchain, V., Hester, J., & Pagès, H. (n.d.).
*SummarizedExperiment: A container (S4 class) for matrix-like assays*
[Bioconductor package].
<https://bioconductor.org/packages/SummarizedExperiment>

Pepe, M. S., Longton, G., Anderson, G. L., & Schummer, M. (2003).
Selecting differentially expressed genes from microarray experiments.
*Biometrics*, *59*(1), 133–142.
<https://doi.org/10.1111/1541-0420.00016>

Thompson, M. L., & Zucchini, W. (1989). On the statistical analysis of
ROC curves. *Statistics in Medicine*, *8*(10), 1277–1290.
<https://doi.org/10.1002/sim.4780081011>

Vivo, J.-M., Franco, M., & Vicari, D. (2018). Rethinking an ROC partial
area index for evaluating the classification performance at a high
specificity range. *Advances in Data Analysis and Classification*,
*12*(3), 683–704. <https://doi.org/10.1007/s11634-017-0295-9>
